---
title: "Models w/o Sedentary Score"
author: "Aine O'Connor"
date: "11/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

# Packages used in analysis
library(haven)
library(sas7bdat)
library(plyr)
library(readr)
library(tidyverse)
library(nnet)
library(MASS)
library(funModeling)
library(brant)
library(arsenal)
library(lmtest)
library(table1)
library(mice)
library(summarytools)
library(car)

# Import and read FLASHE data for adolescent physical activity and demographic

# Read in data: adolescent physical activity
adolescentPA <- read_sas("Teen Physical Activity SAS/teenpa_public_updated.sas7bdat", 
    NULL)

# Read in data: adolescent demographics
adolescentdemo <- read_sas("Teen Demographic SAS/TeenDemo_public_09022015.sas7bdat", NULL)

# Read in data: parent demographics
parentdemo <- read_sas("Parent Demographic SAS/ParentDemo_public_09022015.sas7bdat", NULL)

# Merge datasets by Dyad ID: adolescent physical activity & demographics (yields n=1,661)
adolescent.merged <- merge(adolescentPA, adolescentdemo, by = "DYADID")

# Merge full adolescent dataset with parent dataset by Dyad ID (yields n=1,649)
merged.data <- merge(adolescent.merged, parentdemo, by = "DYADID")

# Create subset of data with variables of interest
df <- data.frame(merged.data$PID.x, merged.data$DYADID, merged.data$XTBMI, merged.data$XTWGHTSTAT, merged.data$XTYAPOUTOFSCHOOL, merged.data$XTYAPWEEKEND, merged.data$XTPREDWEEKOFS, merged.data$XTPREDWEEKW, merged.data$TPANTRAILS, merged.data$TPANPLAYFLD, merged.data$TPANPUBPRK, merged.data$TPANSTORE, merged.data$TPCRIME, merged.data$TPTRAFFIC, merged.data$TAGE, merged.data$TSEX, merged.data$TETHRAC_RC, merged.data$PHSEHLDINCM_RC, merged.data$XPTCOVTYPE)

# Rename variables

# Check variable names
colnames(df)

# Name variables for comprehension
names(df)[1:19]<-c("ID", "DYAD_ID", "BMI", "weight_cat", "PAscore_oos", "PAscore_wknd", "PAmin_oostotal", "PAmin_wkndtotal", "trails", "fields_courts", "parks", "walk_store", "walk_crime", "walk_traffic", "age", "sex", "race", "income", "ins_type")

# Set all -9 values to NA
library(dplyr)
df <- df %>% na_if(., -9)

# Drop NA variables for final dataset (yields n=1,301)
library(tidyr)
df.nomissing <- drop_na(df)

# Compare original merged dataset with the final to see differences
summary(comparedf(df, df.nomissing))

# Look at total missing rate
MissingRate <-  1- nrow(df.nomissing)/nrow(df)

# Change dataset name for ease
data <- df.nomissing

md.pattern(df, plot = TRUE, rotate.names = TRUE)

```
```{r}

# Check variable types
lapply(data, class)

# All variable types are listed as numeric
# Change weight_cat, trails, fields_courts, parks, sex, race, income, ins_type into factors

# Weight
data$weight_catf <- factor(data$weight_cat,
      levels = c(1,2,3,4),
      labels = c("Underweight", "Healthy Weight", "Overweight", "Obese"))
# Check recategorization
table(data$weight_cat, data$weight_catf)

# Trails
data$trailsf <- factor(data$trails,
      levels = c(0,1),
      labels = c("No Access", "Access"))
# Check recategorization
table(data$trails, data$trailsf)

# Fields and courts
data$fields_courtsf <- factor(data$fields_courts,
      levels = c(0,1),
      labels = c("No Access", "Access"))
# Check recategorization
table(data$fields_courts, data$fields_courtsf)

# Parks
data$parksf <- factor(data$parks,
      levels = c(0,1),
      labels = c("No Access", "Access"))
# Check recategorization
table(data$parks, data$parksf)

# Sum outdoor rec variables to produce new outdoor recreation access score
data <- data %>%
  mutate(outdoor_rec = trails + parks + fields_courts)

# Sex
data$sexf <- factor(data$sex,
      levels = c(1,2),
      labels = c("Male", "Female"))
# Check recategorization
table(data$sex, data$sexf)

# Race
data$racef <- factor(data$race,
      levels = c(1, 2, 3, 4),
      labels = c("Hispanic", "Black or African-American Only", "White Only", "Other"))
# Re-level race to set White Only as reference
data$racef <- relevel(data$racef, ref = "White Only")
# Check recategorization
table(data$race, data$racef)

# Income
data$incomef <- factor(data$income,
      levels = c(1,2),
      labels = c("$0 - $99,999", "$100,000+"))
# Check recategorization
table(data$income, data$incomef)

# Insurance
data$ins_typef <- factor(data$ins_type,
      levels = c(1, 2, 3),
      labels = c("Medicaid/SCHIP", "Private Insurance", "Uninsured"))
# Re-level insurance status to set Private Insurance as reference
data$ins_typef <- relevel(data$ins_typef, ref = "Private Insurance")
# Check recategorization
table(data$ins_type, data$ins_typef)

# Change neighborhood walkability & safety variables, which are Likert responses (walk_store, walk_crime, walk_traffic), to binary variables and regroup to indicate walkability/non-walkability

# Walkability to stores
data$walk_store[
    data$walk_store==1|
    data$walk_store==2] <- 0

data$walk_store[
    data$walk_store==3|
    data$walk_store==4] <- 1

# Crime decreases walkability
data$walk_crime[
    data$walk_crime==1|
    data$walk_crime==2] <- 1

data$walk_crime[
    data$walk_crime==3|
    data$walk_crime==4] <- 0

# Traffic decreases walkability
data$walk_traffic[
    data$walk_traffic==1|
    data$walk_traffic==2] <- 1

data$walk_traffic[
    data$walk_traffic==3|
    data$walk_traffic==4] <- 0

# Sum walkability variables to produce new walkability score
data <- data %>%
  mutate(walkability = walk_store + walk_crime + walk_traffic)

# Change age variable values
data$age[data$age==1] <- 11
data$age[data$age==2] <- 12
data$age[data$age==3] <- 13
data$age[data$age==4] <- 14
data$age[data$age==5] <- 15
data$age[data$age==6] <- 16
data$age[data$age==7] <- 17
data$age[data$age==8] <- 18

# Sum minutes of PA variables to produce total MVPA
data <- data %>%
  mutate(PAmin_weeklytotal = PAmin_oostotal + PAmin_wkndtotal)

# Create half-hour measure of PA
data <- data %>%
  mutate(PA_halfhour = PAmin_weeklytotal/30)

# Create variable to indicate meeting/not meeting weekly MVPA totals (~420 mins/wk)
data$PA_meetrec[data$PAmin_weeklytotal < 420] <- "0"
data$PA_meetrec[data$PAmin_weeklytotal >= 420] <- "1"
data$PA_meetrec <- factor(data$PA_meetrec,
      levels = c(0,1),
      labels = c("Does Not Meet Recommendations", "Meets Recommendations"))
class(data$PA_meetrec)

summary(data)

```

### Descriptive statistics

```{r}

# BMI 
descr(data$BMI)
descr(data$PAmin_weeklytotal)

# Relationship between race & outdoor recreation access
ctable(data$trailsf, data$racef, chisq = TRUE)
ctable(data$parksf, data$racef, chisq = TRUE)
ctable(data$fields_courtsf, data$racef, chisq = TRUE)

# Label variables
label(data$age) <- "Age (years)"
label(data$sexf) <- "Sex"
label(data$racef) <- "Race"
label(data$ins_typef) <- "Insurance type"
label(data$walkability) <- "Neighborhood walkability score"
label(data$outdoor_rec) <- "Outdoor recreation access score"
label(data$PAmin_weeklytotal) <- "Total physical activity (mins/wk)"

# Table 1 descriptives
Table1 <- table1(~BMI + racef  + sexf + ins_typef  + walkability + outdoor_rec + PA_meetrec + PAmin_weeklytotal|weight_catf, overall="Total", footnote='SD = standard deviation', data)

Table1

```

### Unadjusted models of DV & IV

```{r}

model1_PAhalf <- lm(BMI ~ PA_halfhour, data=data)
summary(model1_PAhalf)

```

### Partially adjusted model with IV, DV, and confounders

```{r}

model2_PAhalf <- lm(BMI ~ PA_halfhour + age + sexf + racef + ins_typef, data=data)
summary(model2_PAhalf)

```

### Diagnostics

```{r}

# Multicollinearity (VIF < 5)
vif(model2_PAhalf)

# Residual values are independent (Durbin Watson)
dwtest(model2_PAhalf)

# Homoscedasticity (standard predicted values x standardized residuals) and normality of residuals
par(mfrow = c(2,2))
plot(model2_PAhalf)

# Influential cases (?)

```
### Full model

```{r}

model3_PAhalf <- lm(BMI ~ PA_halfhour + age + sexf + racef + ins_typef + outdoor_rec + walkability, data=data)
summary(model3_PAhalf)

```

### Testing for interaction with outdoor recreation

```{r}

model4_outdoorint <- lm(BMI ~ PA_halfhour + age + sexf + racef + ins_typef + outdoor_rec + outdoor_rec*PA_halfhour, data=data)
summary(model4_outdoorint)

```
### Test interaction effect with walkability

```{r}

model5_walkabilityint <- lm(BMI ~ PA_halfhour + age + sexf + racef + ins_typef + walkability +  walkability*PA_halfhour, data=data)
summary(model5_walkabilityint)

```

### Test correlation of walkability and outdoor recreation access

```{r}

histogram_outdoor_rec <- ggplot(aes(x = outdoor_rec), data = data) +
  geom_histogram(binwidth = 1)
histogram_outdoor_rec

histogram_walkability <- ggplot(aes(x = walkability), data = data) +
  geom_histogram(binwidth = 1)

histogram_walkability

cor.test(data$outdoor_rec, data$walkability, method=c("pearson"))

# These variables are highly correlated; walkability is not a true confounder, but is related to outdoor recreation access and should not be treated as a confounder--so removed walkability as confounder

```
### Disaggregating the outdoor rec score

```{r}

model3_rectype <- lm(BMI ~ PA_halfhour + age + sexf + racef + ins_typef + trailsf + parksf + fields_courtsf, data=data)
summary(model3_rectype)

```


### Modelling using binary PA recommendations variable
```{r}

model2_PArec <- lm(BMI ~ PA_meetrec + age + sexf + racef + ins_typef, data=data)
summary(model2_PArec)

model3_PArec <- lm(BMI ~ PA_meetrec + age + sexf + racef + ins_typef + outdoor_rec, data=data)
summary(model3_PArec)

```